% Write in only the text of your abstract, all the extra heading jargon is automatically taken care of
\chapter*{Abstract}
\noindent\justify
This thesis presents searches for physics beyond the Standard Model (SM) of particle physics. 
The searches are performed using $35.9\,\mathrm{fb^{-1}}$ of proton--proton collision data collected with the CMS detector during 2016 at a center-of-mass energy of $\sqrt{s}=13\TeV$, at the CERN Large Hadron Collider (LHC). 
The new physics phenomena targeted assume strong and electroweak production of Supersymmetric (SUSY) particles in the framework of mimimally supersymmetric standard model. 
A powerful tool for searching for SUSY is by using final state scenarios containing two electrons or two muons of opposite electric charge.  
\newpara
\noindent\justify
The document contains a summary of the theoretical framework of the SM and SUSY, along with a comprehensive description of the CMS experiment at the LHC accelerator complex. 
This thesis is based on two analyses, the first targeting strong and electroweak production of superpartners that lead to the leptonic final state along with many jets and large missing transverse momentum. 
The other analysis is targeting the direct production of the superpartners of the SM electrons and muon, using a final state of only leptons and large missing transverse momentum. 
Common to these searches is the use of the missing transverse momentum, which is a crucial variable for searches for $R$-parity conserving SUSY. 
\newpara
\noindent\justify
The thesis contains a chapter on the performance of the missing transverse momentum reconstruction algorithm.  
The remainder of the thesis contains an overview of the search strategies and the estimation of the SM background processes specific to each SUSY production mode. 
%The search for Charginos and Neutralinos result in final states where two or more jets resulting from hadronization are produced, while the search for the direct production of sleptons is chrarcterized by the fact that no hadronization is expected, and thus results in a final state without any jets. The search strategies thus differ slightly, and the two strategies are presented, along with a description of the SM background processes that govern these final states. 
\newpara
\noindent\justify
Since no excess of collision data is observed with respect to the predicted SM backgrounds in neither of the searches, a statistical interpretation of the results is performed to set upper limits in the production cross sections on the SUSY particles. 
These limits greatly extend the limits set using 8\TeV collision data during the LHC Run 1. 
%Concluding remarks commenting on the current absence of evidence for physics beyond the SM (BSM) are given, and an outlook highlighting the unprecedented instantaneous luminosity expected at the LHC, and the window of opportunity for searches for BSM physics that it presents. 
