\chapter*{Acknowledgements}
\noindent\justify
A great number of people deserve recognition for directly or indirectly providing me the support needed to complete this thesis. 
First and foremost, I am grateful for the mentorship of my advisor at ETH Z\"{u}rich, Professor Rainer Wallny. 
I would like to thank Rainer for his support and for all the opportunities he has provided me during the past four years, and for pairing me up with the best post-doc I could ask for, Pablo Martinez Ruiz Del Arbol.
Pablo has provided me the guidance and creativity needed to become a great physicist. 
I consider myself extremely fortunate to have learnt from someone as knowledgable as Pablo, and his constant support and inspiration has improved many workdays. 
I owe an immense amount of gratitude to Zeynep Demiragli, for her leadership, her mentorship, and later on, also our friendship. 
There has been times during my PhD when I lacked motivation and energy. 
Without the tireless support and encouragement of Lesya Shchutska, Loukas Gouskos and Mariarosaria D'Alfonso the MET paper and Slepton paper would not have been published.  
\newpara
\noindent\justify
I was lucky to be surrounded by a great group of friends during my time in Geneva. 
I would especially like to thank Rickard, Sophie, Linnea and Andrew for all the great adventures, R1 breaks and late nights. 
This place would not have been the same without you, and I will cherish our friendships for life. 
To Stany, you are a great inspiration and have always been the best support. I am proud to call you my friend. 
\newpara
\noindent\justify
Finally, I want to express my endless gratitude to my family.
To my older brother, Adam, I am thankful for your brotherly advice and your love. 
I know I always have your support, even though I might not always follow the path you would have wanted. 
To his wife, Linn\'{e}a, I am thankful for our common passion for science and our long friendship. 
Your energy and motivation has always been an inspiration to me, and I am looking forward to having a third doctor in the family :).
To my father, Martin, the first doctor in the family, I am thankful that you early on sparked my interest for science. 
Even though I did not become the doctor you had intended me to, I know how extremely proud you are of me today. 
To my mother, Lillemor, I am thankful for your unconditional love and the close relationship we have had for 30 years, and will have for many more. 
Our weekly phonecalls have been the greatest support I could imagine during these past four years. 
I dedicate this work to my parents, as their undivided love and encouragement is the reason I can call myself a doctor today. 
